Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-07 paper/paper.tex
Edits

\subsubsection{Feature-space Enrichment} ~\\
\textit{Stopword removal}: We used Mallet's TokenSequenceRemoveStopwords Pipe to remove 500 of the most common words in the English language from each tweet. We remove these words because of the high likelihood of many tweets containing these words, without contributing significant discriminatory features to the tweets. %TODO [REFERENCE NLP PAPER ABOUT STOPWORD REMOVAL]
\\
~\\
\textit{Bigrams}: Using n-grams is a common method of increasing the number of features present in short text %[FIND CITATION]
. Often, this is enough to increase accuracy significantly%CITATION
. As discussed in the tokenization section below, we used selective bigramming, which yielded increases in accuracy in almost all test cases.\\
~\\
\textit{HTML corrections}: When accessing tweets through Twitter's public API, the native JSON format includes HTML-encoded entities that can add confusing features to the tweets if not un-escaped. For instance, angle brackets are replaced with \&lt; and \&gt; (less-than and greater-than), causing emoticons, for instance, to no longer be parseable. The emoticon `\textgreater.\textless', for instance, would be `\&gt;.\&lt;'. Necessarily, this Pipe un-escapes all such entities, making them easier to parse.\\
~\\
\textit{URL replacement}: In our dataset, 60\% of the 100,000 tweets we collected contained a link. As the vast majority of these are unique, and shortened using a URL shortener, often Twitter's own t.co. Links, then, lose any features that may have been parsed from the URL, which often contains a domain name or title that may be useful in discriminating especially ambiguous links. Consider, for instance, the tweet `hmmm http://t.co/39D8I3qM' which leaves a parser completely unaware of the true content of the tweet. Following this URL leads to `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade/', which is clearly related to Apple. This pipe replaces the link itself in the tweet with the title of the destination web page: `iPhone 5 vs iPhone 4S: Should You Upgrade? | Redmond Pie.' In cases where the title is empty, we instead replace the shortened link with the resolved URL, which is `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade' in this case. In many cases like this, there is not other way to distinguish the content of a tweet. From our dataset of 100,000 tweets, 102 tweets contained \textit{only} a link, while 608 tweets contained only a link and fewer than 20 other characters. While this is a fairly small portion of the overall dataset for training purposes, it can be crucial in making decision on novel test instances. Pak and Paroubek also noted the need to handle URLs in tweets, however, rather than adding any new text, they chose to remove the link entirely  (\citeauthor{Alexander Pak and Patrick Paroubek}).
~\\
\textit{Spell checking}: Twitter's user-generated content contains a large volume of misspelled words (\citeauthor{Laboreiro:2010:TMM:1871840.1871853}). Using a spell-checking library, we wrote a Pipe to replace unrecognized words (misspelled words) with their corrections if the confidence for correction was above a certain threshold. Our intention was to reduce the noise in the feature set by eliminating common misspellings. This did not provide any statistically significant change in our results. \\
~\\
\subsubsection{Disambiguation}
~\\
\textit{Stemming}: Stemming is the process of reducing words to their `stem,' in order to group different forms of the same word into a single feature. While this stem does not necessarily have to be the root of the word, it should satisfy the property that related words have the same stem. For instance, the stem of the related words `argue,' `argued,' and `arguing' all have the stem of `argu,' even though argu is not the true linguistic stem, or a real word. We used the standard Porter stemmer to implement this (\citeauthor{Porter, M.}).
In some cases, stemming proved to increase our accuracy.
\\

@@

As presented in our results, because of irregular capitalization, two tokens likely to be semantically identical with respect to company classification may be represented in multiple cases. {Nooo, NOOO, nooo, NoOo} is a prime example. To alleviate this, we attempt three different approaches. In the first, we simply lowercase all remaining tokens (note that many URLs become invalid after modifying case). In the second, we lowercase any text that is not in all caps. This serves to retain acronyms and any text specifically place in all caps, while combining words capitalized because of sentence boundaries with their lowercase appearances, for instance. In the third approach, we change all remaining tokens to lowercase, except for leaving the case of all strings containing `apple' (case insensitive) alone. We made this decision based on the semantic difference between the capitalization of Apple the company and apple the fruit, which is a very significant feature in our dataset. While correct capitalization is naturally not always present for this token, this is one case where capitalization serves to disambiguate. In our dataset, `Apple' appeared 11784 times in 10500 tweets, `apple' appeared 4765 times in 4508 tweets, and other case-variations appeared 292 times in 237 tweets (all but one of which was `APPLE'). This same principle can be generalized to a larger set of keywords and applied to other companies in a similar fashion. \\

@@

\section{Experimental Setup}
In order to determine which combination of preprocessing steps would results in the highest performance, we built a custom evaluation platform that iterates through all possible valid combinations of pipes and evaluates their performance. This led to nearly one hundred combinations, as each of the six tokenizer variations was run with and without bigramming, with and without stemming, etc. For consistency, the disjoint training and test sets used to evaluate each combination were identical for each iteration. To best observe the effects of training set size, training sizes of 50 to 1500 were tested in increments of 50, where each set is a superset of the previous set with 50 new randomly chosen instances, as opposed to using n-fold cross validation, which would not display this characteristic evenly across varied training and testing sizes. To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (`apple' or `none' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.

>>>>

\subsubsection{Feature-space Enrichment} ~\\
\textit{Stopword removal}: We used Mallet's TokenSequenceRemoveStopwords Pipe to remove 500 of the most common words in the English language from each tweet. We remove these words because of the high likelihood of many tweets containing these words, without contributing significant discriminatory features to the tweets. %TODO [REFERENCE NLP PAPER ABOUT STOPWORD REMOVAL]
\\
~\\
\textit{Bigrams}: Using n-grams is a common method of increasing the number of features present in short text%[FIND CITATION]
. Often, this is enough to increase accuracy significantly%CITATION
. As discussed in the tokenization section below, we used selective bigramming, which yielded increases in accuracy in almost all test cases.\\
~\\
\textit{HTML corrections}: When accessing tweets through Twitter's public API, the native JSON format includes HTML-encoded entities that can add confusing features to the tweets if not un-escaped. For instance, angle brackets are replaced with \&lt; and \&gt; (less-than and greater-than), causing emoticons, for instance, to no longer be parseable. The emoticon `\textgreater.\textless', for instance, would be `\&gt;.\&lt;'. Necessarily, this Pipe un-escapes all such entities, making them easier to parse.\\
~\\
\textit{URL replacement}: In our dataset, 60\% of the 100,000 tweets we collected contained a link. As the vast majority of these are unique, and shortened using a URL shortener, often Twitter's own t.co. Links, then, lose any features that may have been parsed from the URL, which often contains a domain name or title that may be useful in discriminating especially ambiguous links. Consider, for instance, the tweet `hmmm http://t.co/39D8I3qM' which leaves a parser completely unaware of the true content of the tweet. Following this URL leads to `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade/', which is clearly related to Apple. This pipe replaces the link itself in the tweet with the title of the destination web page: `iPhone 5 vs iPhone 4S: Should You Upgrade? | Redmond Pie.' In cases where the title is empty, we instead replace the shortened link with the resolved URL, which is `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade' in this case. In many cases like this, there is not other way to distinguish the content of a tweet. From our dataset of 100,000 tweets, 102 tweets contained \textit{only} a link, while 608 tweets contained only a link and fewer than 20 other characters. While this is a fairly small portion of the overall dataset for training purposes, it can be crucial in making decision on novel test instances. Pak and Paroubek also noted the need to handle URLs in tweets, however, rather than adding any new text, they chose to remove the link entirely  (\citeauthor{PAK10.385}).
~\\
\textit{Spell checking}: Twitter's user-generated content contains a large volume of misspelled words (\citeauthor{Laboreiro:2010:TMM:1871840.1871853}). Using a spell-checking library, we wrote a Pipe to replace unrecognized words (misspelled words) with their corrections if the confidence for correction was above a certain threshold. Our intention was to reduce the noise in the feature set by eliminating common misspellings. This did not provide any statistically significant change in our results. \\
~\\
\subsubsection{Disambiguation}
~\\
\textit{Stemming}: Stemming is the process of reducing words to their `stem,' in order to group different forms of the same word into a single feature. While this stem does not necessarily have to be the root of the word, it should satisfy the property that related words have the same stem. For instance, the stem of the related words `argue,' `argued,' and `arguing' all have the stem of `argu,' even though argu is not the true linguistic stem, or a real word. We used the standard Porter stemmer to implement this (\citeauthor{porter_1980}).
In some cases, stemming proved to increase our accuracy.
\\

@@

As presented in our results, because of irregular capitalization, two tokens likely to be semantically identical with respect to company classification may be represented in multiple cases. {Nooo, NOOO, nooo, NoOo} is a prime example. To alleviate this, we attempt three different approaches. In the first, we simply lowercase all remaining tokens (note that many URLs become invalid after modifying case). In the second, we lowercase any text that is not in all caps. This serves to retain acronyms and any text specifically placed in all caps, while combining words capitalized because of sentence boundaries with their lowercase appearances, for instance. In the third approach, we change all remaining tokens to lowercase, except for leaving the case of all strings containing `apple' (case insensitive) alone. We made this decision based on the semantic difference between the capitalization of Apple the company and apple the fruit, which is a very significant feature in our dataset. While correct capitalization is naturally not always present for this token, this is one case where capitalization serves to disambiguate. In our dataset, `Apple' appeared 11784 times in 10500 tweets, `apple' appeared 4765 times in 4508 tweets, and other case-variations appeared 292 times in 237 tweets (all but one of which was `APPLE'). This same principle can be generalized to a larger set of keywords and applied to other companies in a similar fashion. \\

@@

\section{Experimental Setup}
In order to determine which combination of preprocessing steps would result in the highest performance, we built a custom evaluation platform that iterates through all possible valid combinations of pipes and evaluates their performance. This led to nearly one hundred combinations, as each of the six tokenizer variations was run with and without bigramming, with and without stemming, etc. For consistency, the disjoint training and test sets used to evaluate each combination were identical for each iteration. To best observe the effects of training set size, training sizes of 50 to 1500 were tested in increments of 50, where each set is a superset of the previous set with 50 new randomly chosen instances, as opposed to using n-fold cross validation, which would not display this characteristic evenly across varied training and testing sizes. To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (`apple' or `none' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.
